Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data

### 2019-11-21 ms/review-responses.tex

\begin{point}{}
    Can the authors provide a bit more intuition behind the patterns of variation seen in generation time, census population size, and variance in the number of offspring with respect to neighborhood size seen in Figure 2? For example, it is not obvious to me why the census population size, for example, should decline systematically with respect to neighborhood size. Presumably this isn't just due to the local demographic stochasticity. Could the authors briefly interpret the observed patterns or cite appropriate literature?
\end{point}

\reply{
    We agree that this is hard to intuit: there are a lot of factors at work.
    We have added some more discussion of these phenomena \llname{censusscaling}.
}

\begin{point}{}
    \autoref{fig:gwas}D: I am surprised by the extent to which the observed values of $-\log_{10}(p)$ fall below the $y=x$ line. Particularly in the lower right panel for large neighbourhood sizes. I would expect that to be close to panmictic - why are the P-values underdispersed? That seems like a potential bug, or else something weird is going on.
\end{point}

\reply{
    We have checked the code to the best of our abilities and did not find a bug causing the underdispersion, but we have switched our visualization in Figure 8D to now show a subset of qqplots as lines, rather than the hexagonal binning over all simulations. This lets us see the density in addition to the range of values, and shows that relatively few points fall below the 1:1 lines. Those that do seem to reflect overcorrection in the regression when using PC coordinates as covariates -- the PCA is capturing some information about the spatial genetic variation which itself covaries only weakly with the phenotype, and as a result we see anomalously low $-log10(p)$ when regressing genotype against phenotype. 
}

@@

    We definately think that a comparison to the spatial Lambda-Fleming-Viot
    would be an extremely useful thing to do, for the reasons the reviewer mentions,
    but this would constitute an entirely different study.
    The SLFV model is only known to be an approximation of one particular biologically explicit model
    (of patchy extinction-recolonization)
    and is only conjectured to be a good approximation
    for spatial models more generally.
    To reiterate, the definition of the SLFV was explicitly motivated by a desire to approximate
    spatial models, especially Bolker-Pacala models like ours.
    Therefore, our study is an important prerequisite
    for simulation study of the spatial Lambda-Fleming-Viot,
    as it works through some of the issues in simulating biologically realistic spatial models,
    and describes some of the patterns.

\begin{point}{\autoref{fig:genparams}:}
    I do not understand why the neighborhood size varies to the same extent in the random mating model as it does for the spatial model. For the random mating model, I would have expected the neighborhood size to be equal to the census size since all individuals have the same probability of being a parent of any given offspring. From [the paragraph at \revref], it is clear that the spatial model would converge to the random mating model when the mean parent-offspring distance tends to infinity only if we were to ignore the impact of range edges. I am thus wondering whether the variation of neighborhood size one observes in Fig 2 for the random mating model is just a consequence of border effects. If that is the case, then the authors should state it clearly and try to justify it from a biological perspective.
\end{point}

\reply{
    This is indeed part of what's going on.
    We have added some discussion of the census size scaling \llname{censusscaling}.
}

@@

\begin{point}{\revref}
    ``Many more species occur in a middle range of neighborhood sizes between 100 and 1000 - a range in which spatial processes play a minor role in our analyses [...]''. Do the authors think that the spatial processes would still play a minor role when neighborhood sizes exceed 100-1000 if the habitat was larger than that taken in the present simulations? It would also probably be useful to mention that neighborhood sizes given in Table 1 should be compared with extreme caution since the size of the corresponding habitats vary across species. More generally, I suspect that the size of the habitat has a substantial impact on the vast majority of statistics examined in this study. Indeed, the mean parent-offspring distance, which is at the core of the definition of Wright's neighborhood size, is only small or large relative to the size of the habitat.
\end{point}

\reply{
This is a good point. Wright's work \citep{Wright1943} suggests some aspects of genetic variation such as variance in allele frequencies and inbreeding coefficients can be estimated by looking only at what he would later \citep{Wright1946} call ``neighborhood size'', but certainly other aspects like the number of segregating sits will also depend on total landscape size. We now note on \llname{popsizecaveat} that we have evaluated only one landscape size, and have added a sentence to the discussion noting that exploration of these patterns in varying landscape sizes is an important avenue for further research \llname{landscapesizecaveat}.

}

\begin{point}{\revref}
    Please add a reference to Guindon, Guo and Welch (2016). This study clearly shows that population density and dispersal parameters are identifiable and can indeed be estimated in practice under the spatial Lambda-Fleming-Viot model.
\end{point}

\reply{
Done. Thank you for pointing us to this study.  \revref
}

### 2019-11-21 ms/spaceness.tex
line qqplots with associated text edits

Figure \ref{fig:gwas}D shows quantile-quantile plots that show the degree of genome-wide inflation of test statistics in PC-corrected GWAS across all simulations and environmental distributions. 
For clinal environments, $-\log_{10}(p)$ values are most inflated when neighborhood sizes are large, consistent with the pattern observed in the count of significant associations after PC regression. In contrast corner and patchy environments cause the greatest inflation in $-\log_{10}(p)$ at neighborhood sizes less than 100, which likely reflects the inability of PCA to account for fine-scale structure caused by very limited dispersal. Finally, we observed that PC regression appears to overfit to some degree for all phenotype distributions, visible in Figure \ref{fig:gwas}D as points falling below the 1:1 line.

\begin{figure*}[p]
\centering
\includegraphics[width=\textwidth]{figures/gwas_summary_loglog.pdf}
    \caption{Impacts of spatially varying environments and isolation by distance on linear regression GWAS. Simulated quantitative phenotypes are determined only by an individual's location and the spatial distribution of environmental factors. In \textbf{A} we show the phenotypes and locations of sampled individuals under four environmental distributions, with transparency scaled to phenotype. As neighborhood size increases a PCA explains less of the total variation in the data (\textbf{B}). Spatially correlated environmental factors cause false positives at a large proportion of SNPs, which is partially but not entirely corrected by adding the first 10 PC coordinates as covariates (\textbf{C}). Quantile-quantile plots in (\textbf{D}) show inflation of $-\log_{10}(p)$ after PC correction across all simulations and environmental distributions, with colors scaled by the median neighborhood size in each region of q-q space.
}
\label{fig:gwas}
\end{figure*}

>>>>

Figure \ref{fig:gwas}D shows quantile-quantile plots for a subset of simulations that show the degree of genome-wide inflation of test statistics in PC-corrected GWAS across all simulations and environmental distributions. 
For clinal environments, $-\log_{10}(p)$ values are most inflated when neighborhood sizes are large, consistent with the pattern observed in the count of significant associations after PC regression. In contrast corner and patchy environments cause the greatest inflation in $-\log_{10}(p)$ at neighborhood sizes less than 100, which likely reflects the inability of PCA to account for fine-scale structure caused by very limited dispersal. Finally, we observed that PC regression appears to overfit to some degree for all phenotype distributions, visible in Figure \ref{fig:gwas}D as points falling below the 1:1 line.

\begin{figure*}[p]
\centering
\includegraphics[width=\textwidth]{figures/gwas_summary_loglog_qqlines.pdf}
    \caption{Impacts of spatially varying environments and isolation by distance on linear regression GWAS. Simulated quantitative phenotypes are determined only by an individual's location and the spatial distribution of environmental factors. In \textbf{A} we show the phenotypes and locations of sampled individuals under four environmental distributions, with transparency scaled to phenotype. As neighborhood size increases a PCA explains less of the total variation in the data (\textbf{B}). Spatially correlated environmental factors cause false positives at a large proportion of SNPs, which is partially but not entirely corrected by adding the first 10 PC coordinates as covariates (\textbf{C}). Quantile-quantile plots in (\textbf{D}) show inflation of $-\log_{10}(p)$ after PC correction for simulations with spatially structure environments, with line colors scaled to the neighborhood size of each simulation.
}
\label{fig:gwas}
\end{figure*}
